An introduction to wavelets using R

### 2022-02-12 text/14/wavelets.Rmd

```{r read-ecg-data}
library("rwt")
library("wavelets")
data(ecg)
head(ecg)
tail(ecg)
```

@@

These coefficients represent the "high pass" filter. There is a complementary "low pass" filter that you get by reversing the coefficients and changing the sign of every other value.

Note: To get the math to work on my end, tThe low pass filter provided by the rwt package needs to be negated.

@@

These two filter pairs are known as a quadrature mirror filter because of the reflection. These two filters are orthonormal.

@@

>>>>

```{r read-ecg-data}
library("ggplot2")
library("magrittr")
library("rwt")
library("wavelets")
data(ecg)
head(ecg)
tail(ecg)
```

@@

These coefficients represent the "high pass" filter. There is a complementary "low pass" filter that you get by reversing the coefficients and changing the sign of every other value. The term "high pass" means that it passes high frequencies above a certain cutoff and attenuates low frequencies. The term "low pass" means the opposite.

Note that not all wavelets develop the low pass filter in the way described above. The construction is known as a quadrature mirror filter. There is an attractive simplicity to this approach.

@@

These two filters are orthonormal.

@@

```{r layout}
x <- c(700, 300, 700)
y <- c(700, 300, 300)
m <- c(
  "W1\nh %*% g", 
  "W1\ng %*% h", 
  "W1\nh %*% h"
)
data.frame(x, y, m) %>%
  ggplot(aes(x, y, label=m)) +
  expand_limits(x=c(100, 900), y=c(100, 900)) + 
  geom_segment(x=100, y=100, xend=900, yend=100) +
  geom_segment(x=900, y=100, xend=900, yend=900) +
  geom_segment(x=900, y=900, xend=100, yend=900) +
  geom_segment(x=100, y=900, xend=100, yend=100) +
  geom_segment(x=500, y=100, xend=500, yend=900) +
  geom_segment(x=100, y=500, xend=900, yend=500) +
  theme_void() + 
  geom_text() -> layout_1
layout_1 +
    geom_text(x=300, y=700, label="V1\ng %*% g")
layout_1+
  geom_segment(x=300, y=500, xend=300, yend=900) +
  geom_segment(x=100, y=700, xend=500, yend=700) +
  geom_text(x=400, y=800, label="W2\nh %*% g") + 
  geom_text(x=200, y=600, label="W2\ng %*% h") + 
  geom_text(x=400, y=600, label="W2\nh %*% h") -> layout_2
layout_2 +
  geom_text(x=200, y=800, label="V2\ng %*% g")
layout_2 +
  geom_segment(x=200, y=700, xend=200, yend=900) +
  geom_segment(x=100, y=800, xend=300, yend=800) +
  geom_text(x=250, y=850, label="W3") + 
  geom_text(x=150, y=750, label="W3") + 
  geom_text(x=250, y=750, label="W3") + 
  geom_text(x=150, y=850, label="etc.")
